babette: BEAUti 2, BEAST2 and Tracer for R

### 2018-05-06 response_letter/response_letter.md

> we have now received two very qualified reviews for your manuscript on babette. 
> Overall, both reviews are very positive and consider your software and article worthy of publication in Methods in Ecology and Evolution. I agree with the reviewers' assessment. Nevertheless, before your manuscript can be accepted for publication, I ask you to follow the advice of the two reviewers carefully to address the minor issues raised by them.

Thank you for these nice words. We are happy to inform you we took care of all these issues.

> I agree with reviewer 2 that 
> more flexible node dating would be an worthwhile feature 
> to add to babette but that this addition could be left for future development. 
> Nevertheless I would find it helpful if implemented options for node dating could be described 
> in slightly more detail in the manuscript; 
> e.g. is an age constraint on the root the only option for node dating, 
> and can it be constrained by prior distributions, or only with a fixed age?

While this manuscript was under review, we already added node dating to the package. Additionally, an example how to use this is added to the manuscript,
where we data the ancestor node of three random taxa assuming monophyly.

@@

> Bilderbeek and Etienne have created a useful tool in babette, and have written a concise and informative paper describing what they have done and how it works. For this reason I am recommending acceptance with minor revisions, and would like to thank the authors for their contribution to phylogenetic software and the BEAST universe.

We are happy to hear our work is appreciated.

> Continuing the grand tradition of picking alliterative puns for BEAST program names, 
> the authors have gone with "babette". 
> I guess that this is in reference to one of the maids that cleans up after The Beast in the famous fairy tale. 
> I think the name is logical and clever. 
> However in the older Disney animated versions of the fairy tale the maid is a "sexy maid" stereotype, 
> and the authors seem to have directly copied the babette logo from the animation. 
> The authors should reconsider this, not only because of the notorious litigousness of Disney lawyers, 
> but also because the optics of using a "sexy maid" logo for scientific software is bad, 
> especially in 2018, especially from two male authors. 
> Sorry if this point gets interpreted as not having a sense of humor, but I think it should be addressed.

Thanks for mentioning this issue in such careful words. We thought it would be indeed a good idea
to follow this advice. All logo's (`babette` and the packages it relies on) have been redesigned (by
a girl actually) to be more gender-neutral.

@@

> Page 3, line 29: "creates a posterior" is a bit of a glib description, 
> and could be reworded as "it uses a Bayesian statistical framework to estimate 
> the joint posterior distribution of estimated phylogenies and..."

Agreed. We use your words to improve the accuracy of the original wording.

> Page 4, lines 97-104: I think the authors are under-selling their own product here. 
> Of course babette does not support everything BEAUTi does at the moment, 
> but instead of highlighting all the things it cannot do, maybe focus more on what it can do. 
> For example, elucidate the 3 prior priors that are supported, 
> instead of saying that only 3 of 7 tree priors are not supported. 
> BEAST2 is very dynamic, and new tree priors are added all the time, 
> so it's likely that 7 will not be the correct number in the future. 
> Likewise I imagine babette will be improved upon and that when someone reads the paper in the future, 
> it may support more than 3 tree priors.

Thanks for pointing this out! Upon a re-read, the original wording did sound rather negative.
The new wording specifies what has been done in a way that would still be correct for future readers.

@@

> Section 3, Usage: there should be some indiciation of how to install babette, 
> or where to go to find out how to install it. 
> There seems to be a minor bug in the installation script: 
> when I installed babette using devtools as recommended, a bunch of dependencies were installed, 
> but when I ran "run" it required an R package "testthat" which was not one of those dependencies. 
> Also, is it possible to export the XML without running babette, 
> or to export the trees after babette has finished running? 
> Those would be two useful examples to include in this section.

We added text on how to install `babette` and BEAST2 (to a default location to be found by `babette`) are
added. The missing `testthat` dependency in the DESCRIPTION file (under `Suggests`) was embarassing, which proves that even a rigid testing
scheme cannot catch all mistakes. An early `babette` user already pointed out
this error and it has been fixed. As, indeed, it is possible to only export the XML
normally created by BEAUti, this has been added, as we expect more people to want this feature.

@@

> Page 5, line 114: "run" is an extremely generic name for a function 
> and is likely to result in namespace conflicts with other packages. 
> Maybe there is a good reason to use this name that I am not aware of, 
> but if not, consider a more informative name.

We completely agree. The function has been renamed to `bbt_run`, following the tradition of `tidyverse` packages
like `stringr`.

> Also the "beast2_jar_path" default is hardcoded to "~/Programs/beast2/lib/beast.jar", 
> which will be wrong for everyone except the authors of the program, 
> so I suggest that this argument should be required rather than optional. 
> At least here in "Usage" a value for this argument should be included.

This point was also mentioned by rOpenSci code reviewers and has been
addressed. Using the `rappdirs` package, folder names can be made platform independent.
Since then, `babette` has been tested to build and pass all tests under both Windows and Linux.

@@

> Page 6, lines 151-155: For parameters which are changed only using a scale operator, 
> and not a random walk or uniform operator, values less than 0 are never proposed. 
> So effectively the prior probability density for something like birth rate 
> with an unbounded uniform prior is 0 to infinity, 
> and having a non-zero probability on negative values is not a valid reason to use an exponential prior.

We agree and wrote down that indeed this is an invalid 
reason ('In practice, this [reasoning] does not matter'). As apparently we did not do a good enough job,
we decided to remove the incorrect reasoning altogether and only demonstrate how to use a different
prior distribution.

> Page 7, line 182: Reference to the Bouckaert & Vaughan blog post 
> doesn't seem relevant to this feature of babette. 
> How does babette fix the crown age? 
> Is it by scaling the root node of the tree to that age, 
> and removing all operators that can change the height of the root node? 
> However it is done, adding a brief explanation in the main text would be good.

The reference to the blog post has been removed. As per this comment, we described how `babette`
fixes a phylogeny crown age, by -indeed- removing all operators that can change the height of the root node. We use the past tense here, as we decided to remove this paragraph althogether: the posteriors produced by BEAST2 have remarkably different branch length distributions, for unknown reasons. Using a dated note on the crown does give the regular branch length distributions.

@@

> Page 7, lines 178-184: Fixing the crown age 
> and estimating the clock is an alternative to fixing the clock 
> and estimating the crown age. 
> I would guess that for some studies this is a useful feature. 
> But in either case the same number of parameters are estimated, 
> so I don't think the overall variation across parameters will be reduced. 
> Also "theoretical" experiments is too vague. 
> Instead you could describe these experiments as 
> "where it is unnecessary to accurately model the uncertainty of absolute divergence times".

Agreed. We added this superior wording to the manuscipt, before removing the entire paragraph, as clarified above.

@@

> As a user of both BEAST2 and R, 
> I agree here is a need for an implementation of BEAST2 in R. 
> My review is based solely on the documentation given by the authors, 
> including the youtube videos, as neither link nor package is available 
> and could therefore not be tested by me. 
> The proposed package babette package combines the base features of BEAST2,
> including standard models and priors, and the associates BEAUti and Tracer. 
> My judgement is that babette appears to be a useful tool, 
> especially when numerous replicates with slighly different settings are to be run.

We are happy to hear the reviewer sees the utility to the package. The reason
the link to CRAN did not work, is as it has not yet been accepted by CRAN. `babette` intends
to be submitted to CRAN, but only after an rOpenSci code review. Such a code review will
help the -overly busy- CRAN maintainers judge `babette` for its high-quality code.

> Personally I would like to see 
> node dating, 
> the multispecies coalescent,
> and ancestral state reconstruction implemented in babette,
> but I suppose this is for future versions of the package.

While this manuscript was under review, we added node dating and an example how to do so. As the first reviewer also suggested this feature, we refer to the comments there.

### 2018-05-07 response_letter/response_letter.md

> we have now received two very qualified reviews for your manuscript on babette. 
> Overall, both reviews are very positive and consider your software and article worthy of publication in Methods in Ecology and Evolution. I agree with the reviewers' assessment. Nevertheless, before your manuscript can be accepted for publication, I ask you to follow the advice of the two reviewers carefully to address the minor issues raised by them.

Thank you for these nice words. We are happy to inform you we took care of all these issues.

> I agree with reviewer 2 that 
> more flexible node dating would be an worthwhile feature 
> to add to babette but that this addition could be left for future development. 
> Nevertheless I would find it helpful if implemented options for node dating could be described 
> in slightly more detail in the manuscript; 
> e.g. is an age constraint on the root the only option for node dating, 
> and can it be constrained by prior distributions, or only with a fixed age?

While this manuscript was under review, we already added node dating to the package. Additionally, we added an example how to to the manuscript,
where we date the ancestral node of three random taxa assuming monophyly.

@@

> Also the "beast2_jar_path" default is hardcoded to "~/Programs/beast2/lib/beast.jar", 
> which will be wrong for everyone except the authors of the program, 
> so I suggest that this argument should be required rather than optional. 
> At least here in "Usage" a value for this argument should be included.

This point was also mentioned by rOpenSci code reviewers and has been
addressed. Using the `rappdirs` package, folder names can be made platform independent.
Since then, `babette` has been tested to build and pass all tests under both Windows and Linux.

> Page 6, lines 151-155: For parameters which are changed only using a scale operator, 
> and not a random walk or uniform operator, values less than 0 are never proposed. 
> So effectively the prior probability density for something like birth rate 
> with an unbounded uniform prior is 0 to infinity, 
> and having a non-zero probability on negative values is not a valid reason to use an exponential prior.

We agree and decided to remove the incorrect reasoning altogether and only demonstrate how to use a different
prior distribution.

@@

> Page 7, line 182: Reference to the Bouckaert & Vaughan blog post 
> doesn't seem relevant to this feature of babette. 
> How does babette fix the crown age? 
> Is it by scaling the root node of the tree to that age, 
> and removing all operators that can change the height of the root node? 
> However it is done, adding a brief explanation in the main text would be good.

The reference to the blog post has been removed. Indeed, we described how `babette`
fixes a phylogeny crown age, by -indeed- removing all operators that can change the height of the root node. But, we decided to remove this paragraph althogether: the posteriors produced by BEAST2 with this setting have remarkably different branch length distributions, for unknown reasons. Using a dated node on the crown does give the regular branch length distributions.

@@

> Page 7, lines 178-184: Fixing the crown age 
> and estimating the clock is an alternative to fixing the clock 
> and estimating the crown age. 
> I would guess that for some studies this is a useful feature. 
> But in either case the same number of parameters are estimated, 
> so I don't think the overall variation across parameters will be reduced. 
> Also "theoretical" experiments is too vague. 
> Instead you could describe these experiments as 
> "where it is unnecessary to accurately model the uncertainty of absolute divergence times".

Agreed. We added this superior wording to the manuscipt, but eventuallt removed the entire paragraph, as clarified above.

@@

> Again, I would like to thank Bilderbeek and Etienne for their software and paper. 
> I am sure babette will be used by others researchers in the future, 
> and they will find it a valuable program.
> 
> Sincerely,
> 
> Huw Ogilvie

We would like to thank Huw Ogilvie for his constructive feedback, that improved the
manuscript in ways that will make it more useful for more readers.

@@

# Reviewer: 2

> As a user of both BEAST2 and R, 
> I agree here is a need for an implementation of BEAST2 in R. 
> My review is based solely on the documentation given by the authors, 
> including the youtube videos, as neither link nor package is available 
> and could therefore not be tested by me. 
> The proposed package babette package combines the base features of BEAST2,
> including standard models and priors, and the associates BEAUti and Tracer. 
> My judgement is that babette appears to be a useful tool, 
> especially when numerous replicates with slighly different settings are to be run.

We are happy to hear the reviewer sees the utility to the package. The reason
the link to CRAN did not work, is that it has not yet been accepted by CRAN. We intend to submit `babette` to CRAN, but only after an rOpenSci code review, which ends at May 18th. Such a code review will
help the -overly busy- CRAN maintainers judge `babette` for its high-quality code.

### 2018-05-07 article_complete.tex
EOD
Build article correctly

\begin{abstract}

  \textbf{1. }
    In the field of phylogenetics, 
    BEAST2 is one of the most widely used software tools. 
    It comes with the graphical user interfaces BEAUti 2, DensiTree and Tracer,
    to create BEAST2 configuration files and to interpret BEAST2's output files. 
    However, when many different alignments or model 
    setups are required, a workflow of graphical user interfaces is cumbersome. \\
  \textbf{2. }
    Here, we present a free, libre and open-source package, \verb;babette;: 
    'BEAUti 2, BEAST2 and Tracer for R', for the R programming language. 
    \verb;babette; creates BEAST2 input files, runs BEAST2 and parses its results, 
    all from an R function call. \\
  \textbf{3. }
    We describe \verb;babette;'s usage and the novel functionality it provides
    compared to the original tools and we give some examples. \\
  \textbf{4. }
    As \verb;babette; is designed to be of high quality and extendable, 
    we conclude by describing the further development of the package. \\
\end{abstract}

{\bf Keywords:} computational biology, evolution, phylogenetics, BEAST2, R

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Phylogenies are commonly used to explore evolutionary hypotheses.
Not only can phylogenies show us how species (or other
evolutionary units) are related to each other, 
but we also estimate relevant parameters such as extinction and 
speciation rates.
There are many phylogenetics tools available to obtain an estimate 
of the phylogenetic tree of a given set of species. 
BEAST2 (\cite{bouckaert2014beast}) is one of the most widely used ones. 
It creates a posterior of jointly estimated phylogenies and model parameters, 
from one or more DNA, RNA or amino acid alignments (see figure \ref{fig:workflow} 
for an overview of the workflow). 
It has a graphical and a command-line interface, 
that both need a configuration file containing 
alignments and model parameters.
BEAST2 is bundled with BEAUti 2 (\cite{drummond2012bayesian}) ('BEAUti' from now on), 
a desktop application to create a BEAST2 configuration file.
BEAUti has a user-friendly graphical user interface, with helpful
default settings.
As such, BEAUti is an attractive alternative 
to manual and error-prone editing of BEAST2 configuration files.

However, BEAUti cannot be called from a command-line script.
This implies that when the user 
wants to explore the consequences of various settings, this must be done manually.
This is the managable workflow when using a few alignments and doing a superficial 
analysis of sensitivity of the reconstructed tree to model settings. 
For exploring many trees (for instance from simulations) and for
more thorough sensitivity analysis, one would like to loop through 
multiple (simulated) alignments, nucleotide substitution models, 
clock models and tree priors. 
One such tool to replace BEAUti is \verb;BEASTmasteR; (\cite{beastmaster}),
which focuses on morphological traits and tip-dating, but also 
supports DNA data. \verb;BEASTmasteR;, however, requires hundreds of
lines of R code to setup the BEAST2 model configuration 
and a Microsoft Excel file to specify alignment files.

BEAST2 is also associated with Tracer (\cite{tracer}) and 
DensiTree (\cite{DensiTree}). Both are desktop applications 
to analyze the output of BEAST2, each with a user-friendly graphical user interface. 
Tracer's purpose is to analyze the parameter estimates generated
from a BEAST2 run. It shows, among
others, the effective sample size (ESS) and time series ('the trace', 
hence the name) of each variable in the MCMC run. Both ESS and trace
are needed to assess the strength of the inference. 
DensiTree visualizes the phylogenies of a BEAST2 posterior, with
many options to improve the simultaneous display of many phylogenies.

However, for exploring the output of many BEAST2 runs, one would like a
script to collect all parameters' ESSes, parameter traces and posterior
phylogenies. There is no single package that offers a complete solution,
but examples of R packages that offer a partial solution are rBEAST
(\cite{rBEAST_ratmann}) and RBeast (\cite{RBeast_faria}). RBeast provides some
plotting options and parsing of BEAST2 output files, but the plotting
functions are too specific for general use, whilst the functions for parsing
call those of \verb;babette;. rBEAST was developed to test a 
particular biological hypothesis (\cite{ratmann2016}), 
and hence was not designed for general use.

Here, we present \verb;babette;:
’BEAUti 2, BEAST2 and Tracer for R’, 
which creates BEAST2 (v.2.4.7) configuration files,
runs BEAST2, and analyzes its results,
all from an R function call. This
will save time, tedious mouse clicking and 
reduces the chances of errors in such repetitive actions.
The interface of \verb;babette; mimics the tools it
is based on. This
familiarity helps both beginner and experienced BEAST2 users 
to make the step from those tools to \verb;babette;.
\verb;babette; enables the creation of a single-script 
pipeline from sequence alignments to posterior analysis in R.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Description}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\verb;babette; is written in the R programming language (\cite{R})
and enables the full BEAST2 workflow from an R function call,
in a similar way to what BEAUti, DensiTree and Tracer do.
\verb;babette;'s main function is \verb;run;, which
configures BEAST2, runs it and parses its output. 
\verb;run; needs at least the name of a 
FASTA file containing a DNA alignment. 
The default settings for the other arguments of \verb;run; 
are identical to BEAUti's and BEAST2's default settings.
Per alignment, a site model, clock model and tree prior can be chosen.
Multiple alignments can be used, each with its own (unlinked) site model, 
clock model and tree prior.

\verb;babette; currently has 108 exported functions to set up  
a BEAST2 configuration file. 
\verb;babette; can currently handle a majority of BEAUti use cases.
Because of BEAUti's high number of plugins, 
\verb;babette; uses a software architecture that is designed to be extended.
Furthermore, \verb;babette; has 13 exported functions to run and help run BEAST2.
One function is used to run BEAST2, others
allow the user to check if a BEAST2 configuration file is indeed valid.
Finally, \verb;babette; has 21 exported function to parse the BEAST2 output
files and analyze the created posterior. \verb;babette; gives the
same ESSes and summary statistics as Tracer. The data is formatted
such that it can easily be visualized using \verb;ggplot2; (for a trace,
similar to Tracer) or \verb;phangorn; (\cite{phangorn}) (for 
the phylogenies in a posterior, similar to DensiTree).

Currently, \verb;babette; does not replace all functionality in BEAUti,
as it does not provide 3 out of 7 tree priors, nor does it support RNA
alignments or use of morphological data. The many plug-ins of BEAUti
are not yet supported by \verb;babette;. \verb;babette; does not support
all command-line arguments of BEAST2, does not provide
the more specialized Tracer analysis options, nor is it as feature-rich
in plotting options as DensiTree. Up until now, the \verb;babette; features 
implemented are those requested by users. Further extension of \verb;babette; 
will be based on future user requests.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Usage}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

In R, the functions of a package need to be loaded in the global namespace first:

\begin{lstlisting}[language=R, floatplacement=H]
library(babette)
\end{lstlisting}
BEAUti, and likewise \verb;babette;, needs at least a FASTA filename
to produce a BEAST2 configuration file. 
In BEAUti, this is achieved by loading a FASTA file, 
then saving an output file using a common
save file dialog. After this, BEAST2 needs to be applied to
the created configuration file. It creates multiple files
storing the posterior. These output
files must be parsed by either Tracer of DensiTree.
In \verb;babette;, all this is achieved by:

\begin{lstlisting}[language=R, floatplacement=H]
out <- run(fasta_filenames = "anthus_aco.fas")
\end{lstlisting}
This code will create a (temporary) BEAST2 configuration file,
from the FASTA file with name \verb;anthus_aco.fas; (which
is supplied with the package, from (\cite{VanEls2018})), 
using the same default settings as BEAUti, which are, 
among others, a Jukes-Cantor site model, a strict clock, and a Yule birth tree prior.
\verb;babette; will then execute BEAST2 using that file, and
parses the output. The returned data structure, named \verb;out;, 
is a list of parameter estimates (called \verb;estimates;), posterior 
phylogenies (called \verb;anthus_aco_trees;, named after
the alignment's name) and MCMC operator performance (\verb;operators;).
An example of using a different site model, clock model 
and tree prior is:

\begin{lstlisting}[language=R, floatplacement=H]
out <- run(
  fasta_filenames = "anthus_aco.fas",
  site_models = create_hky_site_model(),
  clock_models = create_rln_clock_model(),
  tree_priors = create_bd_tree_prior()
)
\end{lstlisting}
This code uses an HKY site model, a relaxed log-normal clock model and a 
birth-death tree prior, each with their default settings in BEAUti.
Table \ref{tab:functions} shows an overview of all functions to 
create site models, clock models and tree priors.
Note that the arguments' names \verb;site_models;, \verb;clock_models; 
and \verb;tree_priors; are plural, as each of these
can be (a list of) one or more elements. Each of these arguments must 
have the same number of elements, so that each alignment has its
own site model, clock model and tree prior. 
An example of two alignments, each with its own site model, is:

\begin{lstlisting}[language=R, floatplacement=H]
out <- run(
  fasta_filenames = c(
    "anthus_aco.fas", 
    "anthus_nd2.fas"
  ),
  site_models = list(
    create_tn93_site_model(), 
    create_gtr_site_model()
  )
)
\end{lstlisting}
\verb;babette; also uses the same default prior distributions as BEAUti 
for each of the site models, clock models and tree priors. 
For example, by default, a Yule tree prior assumes that the birth rate 
follows a uniform distribution, 
from minus infinity to plus infinity. 
This assumption implies that negative and positive birth rates are just as likely, 
where a negative birth rate is biologically impossible (note that 
in practice, this usually works out just fine).
One may prefer an exponential distribution instead, 
as this would assume only positive birth rates, 
and makes high birth rates unlikely.
To do this in \verb;babette;:

\begin{lstlisting}[language=R, floatplacement=H]
out <- run(
  fasta_filenames = "anthus_aco.fas",
  tree_priors = create_yule_tree_prior(
    birth_rate_distr = create_exp_distr()    
  )
)
\end{lstlisting}
In this same example, one may specify
the initial shape parameters of the exponential distribution.
In BEAST2's implementation, an exponential distribution 
has one shape parameter: its mean, which can be set to any
value with \verb;BEAUti;. Within \verb;babette;, to set the 
mean value of the exponential distribution to a 
fixed (non-estimated) value, do:

\begin{lstlisting}[language=R, floatplacement=H]
out <- run(
  fasta_filenames = "anthus_aco.fas",
  tree_priors = create_yule_tree_prior(
    birth_rate_distr = create_exp_distr(
      mean = create_mean_param(
        value = 1.0, 
        estimate = FALSE
      )
    )    
  )
)
\end{lstlisting}
Our initial motivation to create \verb;babette; 
was that we wanted to fix the crown age of a phylogeny.
BEAUti assumes that a phylogeny has a crown age that needs to be 
jointly estimated with the phylogeny and other parameters. 
It does not allow for fixing
the crown age. Without \verb;babette;, one needs to manually edit the BEAST2 
configuration file (\cite{fix_starting_tree}), which is tedious and prone to errors. 
Fixing the crown ages is especially useful for theoretical experiments,
as this allows for one less source of variation.
This is how to specify a fixed crown age with \verb;babette;:

\begin{lstlisting}[language=R, floatplacement=H]
out <- run(
  fasta_filenames = "anthus_aco.fas",
  posterior_crown_age = 15
)
\end{lstlisting}
\verb;babette; allows for the same functionality as Tracer.
Tracer works on the values of the parameter estimates sampled
in the BEAST2 run. This is called the "trace" (hence the name).
The start of the trace is usually discarded, as an MCMC 
algorithm (such as used by BEAST2) first has to converge to
its equilibrium. The start of the trace, called the "burn-in", 
will be removed, because its parameter estimates are not 
representative. By default, Tracer discards the first 10\% of all 
the parameter estimates. 
To remove a 20\% burn-in from all parameter estimates 
in \verb;babette;, the following code can be used:

\begin{lstlisting}[language=R, floatplacement=H]
traces <- remove_burn_ins(
  traces = out$estimates, 
  burn_in_fraction = 0.2
)
\end{lstlisting}
Tracer shows the ESSes of each posterior's variables.
These ESSes are important to determine the strength of the
inference. As a rule of thumb, an ESS of 200 is acceptable 
for any parameter estimate.
To calculate the effective sample sizes (of all estimated variables) in \verb;babette;:

\begin{lstlisting}[language=R, floatplacement=H]
esses <- calc_esses(
  traces = traces, 
  sample_interval = 1000
)
\end{lstlisting}
Tracer displays multiple summary statistics for each
estimated variable: the mean and its standard error, standard deviation,
variance, median, mode, geometric mean, 95\% highest posterior density interval, 
auto-correlation time and effective sample size. It displays these statistics per
variable. In \verb;babette;, these summary statistics are collected for
all estimated parameters at once:

\begin{lstlisting}[language=R, floatplacement=H]
sum_stats <- calc_summary_stats(
  traces = traces, 
  sample_interval = 1000
)
\end{lstlisting}
\verb;babette; allows for the same functionality as DensiTree.
DensiTree displays the phylogenies in a posterior at the same
time scale, drawn one over one another, allowing to see the uncertainty in
topology and branch lengths. Within the object \verb;out;, 
the posterior phylogenies are stored as \verb;anthus_aco_trees;,
and can be plotted as such:

\begin{lstlisting}[language=R, floatplacement=H]
plot_densitree(phylos = out$anthus_aco_trees)
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{babette resources}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\verb;babette; is free, libre and open source software available 
from the official R package archive at 
\url{http://cran.r-project.org/src/contrib/PACKAGES.html\#babette}
and is licensed under the GNU General Public License v3.0.
\verb;babette; uses the Travis CI (\url{https://travis-ci.org})
continuous integration service, which is known to significantly 
increase the number of bugs exposed (\cite{vasilescu2015}) and increases
the speed at which new features are added (\cite{vasilescu2015}).
\verb;babette; has a 100\% code coverage, which correlates with 
code quality (\cite{horgan1994,del1995correlation}). 
\verb;babette; follows Hadley Wickham's style guide (\cite{style_guide}), 
which improves software quality (\cite{fang2001}).
\verb;babette; depends on multiple packages, which are 
\verb;ape; (\cite{APE}), 
\verb;beautier; (\cite{beautier}),
\verb;beastier; (\cite{beastier}),
\verb;devtools; (\cite{devtools}),
\verb;geiger; (\cite{GEIGER}),
\verb;ggplot2; (\cite{ggplot2}),
\verb;knitr; (\cite{knitr}),
\verb;phangorn; (\cite{phangorn}),
\verb;rmarkdown; (\cite{rmarkdown}),
\verb;seqinr; (\cite{seqinr}),
\verb;stringr; (\cite{stringr}),
\verb;testit; (\cite{testit}) and 
\verb;tracerer; (\cite{tracerer}).
We tested \verb;babette; to give a clean 
error message for incorrect input, by
calling \verb;babette; one million times
with random or random sensible inputs, 
using the Peregrine high performance computer cluster. 
The scripts to do so are supplied with \verb;babette;.

\verb;babette;'s development takes place on GitHub,
\url{https://github.com/richelbilderbeek/babette}, 
which accommodates collaboration (\cite{perez2016ten}) 
and improves transparency (\cite{gorgolewski2016practical}).
\verb;babette;'s GitHub facilitates feature requests and 
has guidelines how to do so.

\verb;babette;'s documentation is extensive. All functions are documented
in the package's internal documentation. For quick use, 
each exported function shows a minimal example. 
For easy exploration, each exported function's documentation links to related functions.
Additionally, \verb;babette; has a vignette that demonstrates extensively how
to use it. There is documentation on the GitHub to get started, 
with a dozen examples of BEAUti screenshots with equivalent \verb;babette; code.
Finally, \verb;babette; has tutorial videos that can 
be downloaded or viewed on YouTube, \url{https://goo.gl/weKaaU}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Citation of babette}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Scientists using \verb;babette; in a published paper can cite this
article, and/or cite the \verb;babette; package 
directly. To obtain this citation from within an R script, use:

\begin{lstlisting}[language=R]
> citation("babette")
\end{lstlisting}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Thanks to Yacine Ben Chehida and Paul van Els for supplying their 
BEAST2 use cases. Thanks again to Paul van Els for the sharing his FASTA files 
for use by this package. Thanks to Leonel Herrera-Alsina, Raphael Scherrer 
and Giovanni Laudanno for their comments on this package and article.
We would like to thank the Center for Information Technology of the University 
of Groningen for their support and for providing access to the Peregrine 
high performance computing cluster.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Authors' contributions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

RJCB created the package. 
RJCB wrote the first manuscript draft. 
All authors critically revised the text and gave approval for publication.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MEE style
\bibliographystyle{mee}
\bibliography{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[h]
\centering
\begin{tabular}{ | l | l | }
\hline
\textbf{Name} & \textbf{Description} \\
\hline
\verb;run; & Run BEAST2 \\
\hline
\verb;create_gtr_site_model; & Create a GTR site model \\
\verb;create_hky_site_model; & Create an HKY site model \\
\verb;create_jc69_site_model; & Create a Jukes-Cantor site model \\
\verb;create_tn93_site_model; & Create a TN93 site model \\
\hline
\verb;create_rln_clock_model; & Create a relaxed log-normal clock model \\
\verb;create_strict_clock_model; & Create a strict clock model \\
\hline
\verb;create_bd_tree_prior; & Create a birth-death tree prior \\
\verb;create_cbs_tree_prior; & Create a coalescent Bayesian skyline tree prior \\
\verb;create_ccp_tree_prior; & Create a coalescent constant-population tree prior \\
\verb;create_cep_tree_prior; & Create a coalescent exponential-population tree prior \\
\verb;create_yule_tree_prior; & Create a Yule tree prior \\
\hline
\verb;create_beta_distr; & Create a beta distribution \\
\verb;create_exp_distr; & Create an exponential distribution \\
\verb;create_gamma_distr; & Create a gamma distribution \\
\verb;create_inv_gamma_distr; & Create an inverse gamma distribution \\
\verb;create_laplace_distr; & Create a Laplace distribution \\
\verb;create_log_normal_distr; & Create a log-normal distribution \\
\verb;create_normal_distr; & Create a normal distribution \\
\verb;create_one_div_x_distr; & Create a 1/X distribution \\
\verb;create_poisson_distr; & Create a Poisson distribution \\
\verb;create_uniform_distr; & Create a uniform distribution \\
\hline
\end{tabular}
\caption{babette's main functions}
\label{tab:functions}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\end{document}
\grid
\grid
\grid
\grid
\grid

>>>>

@@

\end{document}
